---
title: "Automorphism analysis on COVID-19 data"
author: 
    - name: Robersy Sanchez
      affiliation: Department of Biology.
                   Pennsylvania State University, University Park, PA 16802
      email: rus547@psu.edu
date: "`r format(Sys.time(), '%d %B %Y')`"
fontsize: 11pt
fontfamily: "serif"
output:
    BiocStyle::html_document:
        toc: true
        toc_depth: 3
        toc_float: 
                collapsed: false
                smooth_scroll: true
        number_sections: true
        theme: united 
        geometry: margin=0.8in
        highlight: tango
toc_float: true
toc_depth: 3
abstract: |
    Analysis of COVID-19 mutational events in terms of automorphisms.
vignette: >
    %\VignetteIndexEntry{Get started-with Methyl-IT}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::knitr}
---
    
    <style type="text/css">
    
    .watch-out {
        max-width: 1820px !important;
        margin-left: auto;
        margin-right: auto;
        background-color: #c9feed;
            font-weight: bold
        font-size: 100%;
        padding: 2px;
    }

pre, code {
    max-width: 1820px !important;
    margin-left: auto;
    margin-right: auto;
    background-color: #c9feed;
        font-weight: bold
    font-size: 100px;
    padding: 2px;
}

code {
    font-size: 100%;
    font-family: Consolas,Monaco,"Courier New",monospace;
    border-radius: 0;
}

code.r{
    font-size: 100%;
}
</style>
    
    ```{r, include = FALSE}
knitr::opts_chunk$set(
    class.source = "watch-out",
    collapse = TRUE,
    comment = "#>"
)
```


```{r set-options, echo=FALSE, cache=FALSE}
options(width = 120)
```

```{r lib}
library(Biostrings)
library(GenomAutomorphism)
library(usefr)
```

# SARS coronavirus GZ02 vs bat-SL-CoVZC45 
```{r dat}
URL <- paste0("https://github.com/genomaths/seqalignments/raw/master/", 
        "COVID-19/AY390556.1_265-13398_13398-21485_RNA-POL_SARS_COVI_GZ02.fas")

covid_aln2 <- readDNAMultipleAlignment(filepath = URL)
```


```{r aut_1}
covid_aut <- automorphism(
                    seq = covid_aln2,
                    group = "Z64",
                    cube = c("ACGT", "TGCA"),
                    cube_alt = c("CATG", "GTAC"),
                    verbose = FALSE)

covid_aut
```


```{r barplot, fig.height = 5, fig.width = 6}
counts <- table(covid_aut$cube[ covid_aut$autm != 1 | is.na(covid_aut$autm) ])

par(family = "serif", cex = 0.9, font = 2, mar=c(4,6,4,4))
barplot(counts, main="Automorphism distribution",
        xlab="Genetic-code cube representation",
        ylab="Fixed mutational events",
        col=c("darkblue","red", "darkgreen"), 
        border = NA, axes = FALSE, #ylim = c(0, 2000), 
        cex.lab = 2, cex.main = 1.5, cex.names = 2)
axis(2, at = c(0, 500, 1000, 1500, 2000), cex.axis = 1.5)
mtext(side = 1,line = -1.5, at = c(0.7, 1.9, 3.1, 4.3, 5.5),
      text = paste0( counts ), cex = 1.4,
      col = c("white","yellow", "black"))

```

# Analysis of conserved regions
```{r uniq}
conserv2 <- conserved_regions(covid_aut)
conserv2
```


Next, function [fitCDF](https://rdrr.io/github/genomaths/usefr/man/fitCDF.html)
is applied to find the best fitted probability function to the the sizes of
conserved regions. 
```{r dist, fig.height= 8, fig.width=12}
widths <- width(conserv2)
dist2 <- fitCDF(widths, distNames = c(2, 3, 7, 10, 11, 19, 20), plot = TRUE, 
               loss.fun = "cauchy")
dist2
```


```{r hist, fig.height = 4, fig.width = 6}
par(lwd = 0.5)
hist(widths, 14, freq = FALSE, las = 1, family = "serif",
family = "serif", col = "cyan1", cex.main = 0.9,
main = "Histogram and best fitted CDF model for conserved region sizes",
xlab = "Conserved region size")
x1 <- seq(1, 150, by = 1)
lines(x1, dggamma(x1,
                  alpha = coef(dist2$bestfit)[1],
                  scale = coef(dist2$bestfit)[2],
                  psi = coef(dist2$bestfit)[3]
                  ),
      col = "red", lwd = 1)
```


After apply Bootstrap test for Goodness of fit,  tt seems to be that we have 
no reasons to reject the hypothesis that the sizes of conserved regions follows 
Generalized Gamma distribution.
```{r gof2}
mcgoftest( 
          varobj = widths, 
          model = dist2,
          stat = "ks")

cat("\n===========================\n")
mcgoftest(
          varobj = widths,
          model = dist2,
          stat = "rmse",
          breaks = 12)

```

## Grouping automorphism by automorphism's coefficients. Types of mutations

```{r autby3}
autby_coef2 <- automorphismByCoef(covid_aut)
autby_coef2 <- autby_coef2[ autby_coef2$autm != 1 & autby_coef2$autm != -1  ]
```


Alignment gaps denoting indel mutations are labeled as “---”. 
```{r barplot2, fig.height = 14, fig.width = 35}
counts2 <- table(autby_coef2$mut_type)
counts2 <- sort(counts, decreasing = TRUE)
count. <- counts[ counts > 2 ]

par(family = "serif", cex.axis = 2, font = 2, las = 1, 
    cex.main = 1.4, mar = c(4,1,4,4))
barplot(count., main="Automorphism distribution per Mutation type",
        col = colorRampPalette(c("red", "yellow", "blue"))(36), 
        border = NA, axes = FALSE,las=2)
axis(side = 2,  cex.axis = 2, line = -2.8 )
counts2
```



# SARS coronavirus GZ02 vs bat-SL-CoVZC45 and Bat SARS-like coronavirus isolate Rs7327 

```{r dat2}
data(covid_autm, package = "GenomAutomorphism")
covid_autm
```


```{r conserved_regions}
conserv <- conserved_regions(covid_autm)
conserv
```


```{r consv}
consvr <- c(conserv, conserv2)
widths <- width(consvr)
dist <- fitCDF(widths, distNames = c(2, 3, 7, 10, 11, 19, 20), plot = TRUE, 
               loss.fun = "cauchy")
dist
```

Joining the datasets increases the sample size and improved predictions
```{r hist2, fig.height = 4, fig.width = 6}
par(lwd = 0.5)
hist(widths, 14, freq = FALSE, las = 1, family = "serif",
family = "serif", col = "cyan1", cex.main = 0.9,
main = "Histogram and best fitted CDF model for conserved region sizes",
xlab = "Conserved region size")
x1 <- seq(1, 150, by = 1)
lines(x1, dggamma(x1,
                  alpha = coef(dist$bestfit)[1],
                  scale = coef(dist$bestfit)[2],
                  psi = coef(dist$bestfit)[3]
                  ),
      col = "red", lwd = 1)
```


```{r gof3}
mcgoftest( 
          varobj = widths, 
          model = dist,
          stat = "ks")

cat("\n===========================\n")
mcgoftest(
          varobj = widths,
          model = dist,
          stat = "rmse",
          breaks = 12)

```

## Grouping automorphism by automorphism's coefficients

```{r autby2}
autby_coef <- automorphismByCoef(covid_autm)
autby_coef <- c(autby_coef, autby_coef2)
autby_coef <- autby_coef[ autby_coef$autm != 1 & autby_coef$autm != -1  ]
```



alignment are labeled as “---”. 
```{r barplot_3, fig.height = 14, fig.width = 12}
counts <- table(autby_coef$mut_type)
counts <- sort(counts, decreasing = TRUE)
count. <- counts[ counts > 2 ]

par(family = "serif", cex.axis = 2, font = 2, las = 1, 
    cex.main = 1.4, cex.lab = 2)
barplot(count., main="Automorphism distribution per Mutation type",
        col = colorRampPalette(c("red", "yellow", "blue"))(36), 
        border = NA, axes = TRUE,las=2)
```



